---
title: "Data Tidying"
author: "Mary McCabe"
date: "October 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Download packages to be able to use specific commands.
```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
```

Read in catch data from INSERT CITATION and WEBSITE URL here.
Right click on download link to directly read into to R.

```{r}
catch_df <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)
head(catch_df)
```

%>% Pipe Operator
Computer shortcut is Ctrl + Shft + m


`select` the columns we need for analysis  
`gather` to move from wide to long  
key is associated with column names, value  
`spread` to move from long to wide  
`rename` to rename a column

```{r}
catch_clean <- catch_df %>% 
  #select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
  select(-All, -notesRegCode) %>% 
  gather(key = "species", value = "catch", -Region, -Year) %>% 
  rename(catch_thousands = catch)


head(catch_clean)
```

Testing for the Error

```{r}
test <- as.numeric(catch_clean$catch_thousands)
i <- which(is.na(test) == T)
print(i)
catch_clean[i,]
```

`mutate` to create a new catch column

```{r}
catch_clean_mutated <- catch_clean %>%
  mutate(catch_thousands = ifelse(catch_thousands == "I", 1, catch_thousands)) %>% 
mutate(catch_thousands = as.numeric(catch_thousands)) %>% 
mutate(catch = catch_thousands * 1000)


head(catch_clean_mutated)
```

split using `group_by`

```{r}
catch_summarized <- catch_clean_mutated %>% 
  group_by(Region, Year) %>% 
  summarize(mean_catch = mean(catch),
            n_obs = n())
            

head(catch_summarized)
```

```{r}
catch_chinook <- catch_clean_mutated %>% 
  filter(species == "Chinook") %>% 
  arrange(-catch)
```